---
title: "Youth Risk Behavior Surveillance"
author: "Luca Toraldo"
date: "10/19/2020"
output: html_document
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE, 
                      tidy=FALSE,
                      size="small")
options(digits = 3)
knitr::opts_chunk$set(fig.width=6.75, 
                      fig.height=6.75,
                      fig.align = "center")
```
```{r load-libraries, echo=FALSE}
library(tidyverse)
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
library(scales)
```

# Youth Risk Behavior Surveillance

Every two years, the Centers for Disease Control and Prevention conduct the [Youth Risk Behavior Surveillance System (YRBSS)](https://www.cdc.gov/healthyyouth/data/yrbs/index.htm) survey, where it takes data from high schoolers (9th through 12th grade) to analyze health patterns. We will work with a selected group of variables from a random sample of observations during one of the years the YRBSS was conducted.

## Load the data

This data is part of the `openintro` textbook and we can load and inspect it. There are observations on 13 different variables, some categorical and some numerical.
```{r}
# Loading the YRBSS data set
data(yrbss)

# Examining the variables and observations in the data set
glimpse(yrbss)
```

## Exploratory Data Analysis

We will first start with analyzing the `weight` of participants in kilograms. Using visualization and summary statistics, we will describe the distribution of weights.We are missing 1,004 observations for weight in the data set. 
```{r, eda_on_weight}
yrbss_weights <- yrbss %>% 

  # Filtering for missing values in 'weight'
   filter(!weight %in% NA) %>% 

  # Generating summary statistics
  summarise(mean = mean(weight),
            median = median(weight),
            max = max(weight),
            min = min(weight),
            sd = sd(weight))

# Displaying summary statistics
yrbss_weights

# Creating distribution histogram
ggplot(yrbss, aes(x = weight)) +
  geom_histogram() +
  scale_y_continuous(label = comma) +
  theme_bw() +
  labs(title = "Weighty Children",
       subtitle = "Distribution of Weights for American High Schoolers",
       x = "Weight (kg)",
       y = "Count",
       caption = "Source: CDC")
```

Next, we consider the possible relationship between a high schooler’s weight and their physical activity. Plotting the data is a useful first step because it helps us quickly visualize trends, identify strong associations, and develop research questions.

Let’s create a new variable `physical_3plus`, which will be `yes` if they are physically active for at least 3 days a week, and `no` otherwise.
```{r}
yrbss <- yrbss %>% 

  # Creating 'physical_3plus' variable and adding it to data set
  mutate(physical_3plus = ifelse(physically_active_7d >= 3, "yes", "no"))

yrbss %>% 
  
# Filtering for missing values in 'physical_3plus'
  filter(!is.na(physical_3plus)) %>% 
  
# Finding proportion of high schoolers that do and do not get adequate physical activity  
  group_by(physical_3plus) %>% 
  summarise(count = n()) %>% 
  mutate(prop = count / sum(count))
```
Here is a 95% confidence interval for the population proportion of high schoolers that are *NOT* active 3 or more days per week.
```{r}
# Creating bootstrap simulation to generate confidence interval
weight_ci <- yrbss %>% 
  specify(response = physical_3plus, success = "no") %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "prop") %>% 
  get_confidence_interval(level = 0.95, type = "percentile")
weight_ci
```
  
Below is a boxplot of `physical_3plus` vs. `weight`. There appears to be a weak relationship between these two variables. Both medians are around the 60 to 70 kg range and both categories contain a high number of outliers on the higher end. While this is generally not what would be expected from this information, a possible shortcoming is that 'weight' captures both fat and muscle, meaning that some of the high-end outliers for the 'yes' group could be students that are highly fit and muscular. 
```{r, boxplot}
# Creating boxplot broken down by 'physical_3plus'
weight_box <- yrbss %>% 
  filter(!physical_3plus %in% NA) %>% 
  ggplot(aes(x = weight, y = physical_3plus)) +
  geom_boxplot() +
  theme_bw() +
  labs(title = "Muscle Mass?",
       subtitle = "Distribution of Weights for American High Schoolers by Physical Activity",
       x = "Weight (kg)",
       y = "3+ Hours of Physcical Activity per Day",
       caption = "Source: CDC")
weight_box
```

## Confidence Interval

Boxplots show how the medians of the two distributions compare, but we can also compare the means of the distributions using either a confidence interval or a hypothesis test.
```{r}
yrbss %>%
  group_by(physical_3plus) %>%
  filter(!is.na(physical_3plus)) %>% 
  
# Generating summary statistics that will be used to create confidence interval  
  summarise(mean_weight = mean(weight, na.rm = TRUE),
            sd_weight = sd(weight, na.rm=TRUE),
            count = n(),
            se_weight = sd_weight/sqrt(count),
            t_critical = qt(0.975, count-1), 
            margin_of_error = t_critical * se_weight,
            
# Creating the confidence interval as the spread bewtween 'lower' and 'upper'            
            lower = mean_weight - t_critical * se_weight,
            upper = mean_weight + t_critical * se_weight)
```
There is an observed difference of about 1.77kg (68.44 - 66.67) and we notice that the two confidence intervals do not overlap. It seems that the difference is at least 95% statistically significant. Let us also conduct a hypothesis test.

## Hypothesis test with formula

The null hypothesis for this test is that the true difference in means is equal to 0 while the alternative hypothesis is that the true difference in means is not equal to 0.
```{r}
# Conducting the hypothesis test
t.test(weight ~ physical_3plus, data = yrbss)
```

## Hypothesis test with `infer`

Next, we will introduce a new function, `hypothesize`, that falls into the infer work flow. We will use this method for conducting hypothesis tests. But first, we need to initialize the test, which we will save as `obs_diff`.
```{r}
# Initializing the 'hypothesize' test
obs_diff <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
```
After we have initialized the test, we need to simulate the test on the null distribution, which we will save as null.
```{r}
# Simulating previously initialized test on a null distribution
null_dist <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
```
Here, `hypothesize` is used to set the null hypothesis as a test for independence, i.e., that there is no difference between the two population means. Also, note that the `type` argument within generate is set to permute, which is the argument when generating a null distribution for a hypothesis test.

We can visualize this null distribution with the following code:
```{r}
# Visualizing the null distribution
ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()
```
We can also calculate the p-value for our hypothesis test using the function `infer::get_p_value()`.
```{r}
# Generating visualization of location of p-value
null_dist %>% visualize() +
  shade_p_value(obs_stat = obs_diff, direction = "two-sided")

# Calculating p-value
null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")
```
This the standard work flow for performing hypothesis tests.
